---
title: "Assignment 3"
author: "Filip Mellgren"
date: '2019-02-23'
output: 
  pdf_document:
    df_print: tibble
    highlight: tango
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```

# Theoretical exercises




# Empirical exercises
Do exercises 10a-10g in the textbook (p.340)

- Remark 1: It is possible that the
values you obtain for the F-statistics, p-values and correlations are different than those
reported since the sample is extended. However, the main conclusions should be the same. 

- Remark 2: Exercise d. is optional and so is the part on the forecast error variance in e. (but you could use the command fevd in STATA to answer these questions).

- Remark 3: You find the appropriate specifications for the variables st
, $\Delta$lip, and $\Delta$ur described in the text to exercise 9 (p.339).

```{r}
# Load necessary packages
library(rio)
library(tidyverse)
library(vars)
```


### 10: 
Estimate the three-VAR beginning in 1961Q1 and use the ordering such that $\Delta lip_t$ is causally prior to $\Delta ur_t$ and that $\Delta ur_t$ is causally prior to $s_t$.
```{r}
# Load the data
df <- as_tibble(import("A3_2019_data.dta"))
```

```{r}
# Make the date variable a number
df <- df %>% mutate(year = as.numeric(str_sub(DATE, start = 1, end = 4)),
                Q = as.numeric(str_sub(DATE,-1, -1)),
                DATE = year + (Q-1)/4)

# Need to define some new variables as well
df <- df %>% mutate(dlip = log(indprod)-log(lag(indprod)), 
                    dur = urate - lag(urate),
                    s = r10 - tbill)
```

```{r}
# Estimate the 3 VAR, with the ordering dlip, dur, s
```

### 10 a:
If you perform a test to determine whether $s_t$ Granger causes $\Delta lip_t$, you should find that the F-statistic is 2.44 with a prob-value of 0.065. How do you interpret this result?
```{r}
# might use grangertest()
```

### 10 b:
Verify that $s_t$ Granger causes $\Delta unemp_t$. You should find that the F statistic is 5.93 with a prob value of less than 0.001.
```{r}
# might use grangertest()
```

### 10 c:
It turns out that the correlation coefficient between $e_{1t}$ and $e_{2t}$ is -0.72. The correlation between $e_{1t}$ and $e_{3t}$ is -0.11 and between $e_{2t}$ and $e_{3t}$ is 0.10. Explain why the ordering of a Choleski composition is likely to be important for obtaining the impulse responses.


### 10 e:
Now estimate the model using the levels of $lip_t$ and $ur_t$. Do you now find a lag length of 5 appropriate?

```{r}
df <- df %>% mutate(lip = log(indprod))
```

### 10 f:
Obtain the impulse response function from the model using $\Delta lip_t, \Delta ur_t$ and $s_t$. Show that a positive shock to the industrial production induces a decline in the unemployment rate that lasts six quarters. Then, $\Delta ur_t$ overshoots its long run level before returning to zero.

```{r}

```


### 10 g: 
Reverse the ordering and explain why the results depend on whether or not $\Delta lip_t$ proceeds $\Delta ur_t$

```{r}

```


