---
title: "Assignment 4"
author: "Filip Mellgren, Ismael Moreno Martinez, Athina Swahn"
date: '2019-03-03'
output: pdf_document
---
# Part I

## 1 a 

Rewriting the equation in the following way: 

$$-\alpha_s(r_{Lt-1}-\beta r_{St-1}-\mu)= -\Delta r_{St}+\sum^2_{i=1}a_{i,11}\Delta r_{St-1}+ \Delta a_{i,12}\Delta r_{Lt-i}+\epsilon_{St}$$
It is clear from the equation above that all RHS variables are I(0). Therefore, 
the linear combination $r_{Lt-1}-\beta r_{St-1}$ must be a stationary process. 
That would mean that $r_{St}$ and $r_{Lt} must have a stochastic trend in common. 
Hence, the cointergration vector is \[B=
    \left[ {\begin{array}{cc}
  1  \\
   -\beta \\
  \end{array} } \right]\]
  
The long run equilibrium is defined by $r_{Lt}= \beta r_{St}$

## 1 b 
$r_{Lt}$ does not Granger cause $r_{St}$
$H_0: \Delta a_{i,12}=\alpha_S= 0$

## 1 c 
assuming $\alpha_s , \alpha_L >0$
(i) $r_{Lt-1}>\beta r_{St-1}+ \mu$
(ii) $r_{Lt-1}< \beta r_{St-1}+ \mu$
assuming $\alpha_S >0$ and $\alpha_L = 0$
(iii) $r_{Lt-1} > \beta r_{St-1} + \mu$
(iv) $r_{Lt-1}< \beta r_{St-1}$

## 1d
We can't run the regression because it's a spurious correlation. This means the residuals are serially correlated, and that there is no linear relationship that can mitigate this. Hence, any estimated regression will suffer rom having residuals integrated of order one.


# Part II
Do exercise 4 (but not 4f) in the textbook (pp.402-403). Remark: It is possible that the values you obtain differ from those reported in the text to the exercise since the sample is extended. However, the main conclusions should be the same.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r}
library(rio)
library(tidyverse)
library(urca)
```

```{r}
df <- import("A4_2019.dta") %>% as_tibble()
```

```{r}
# Make the date variable a number
df <- df %>% mutate(year = as.numeric(str_sub(DATE, start = 1, end = 4)),
                Q = as.numeric(str_sub(DATE,-1, -1)),
                DATE = year + (Q-1)/4)
```
Begin by plotting the series:
```{r}

df %>% gather(tbill, r5, r10, key = "duration", value = "rate") %>%
  ggplot(aes(x = DATE, y = rate, color = duration)) +
  geom_line() +
  labs(x = "date", title = "Non stationarity", 
       subtitle = "But potentially a common stochastic trend")
```


## 4a
Pretest to show that all variables act as unit root processes using ADF with lag length equal to the longest lag length with significicant at the 5% level, including an intercept but not a time trend.

```{r}
# Start with a long lag length, then run several times or just use what the
# book found directly. NB, the data set we have is longer so might have
# changed.
tbill_adf_7 <- ur.df(df$tbill, type = "drift", lags = 7)
summary(tbill_adf_7)

r5_adf_5 <- ur.df(df$r5, type = "drift", lags = 7)
summary(r5_adf_5)

r10_adf_7 <- ur.df(df$r10, type = "drift", lags = 5)
summary(r10_adf_7)

# May want to condense into a smaller table simialr to the one in the book

```
The appropriate lag lengths for the extended data set are 7 for the tbill and r5, and 5 for r10, which differs somewhat from the lag lengths in Enders p.402.
Comment, can we reject the null of a unit root? 

## 4b
Use the Engle-Granger procedure to estimate cointergrating relationships.

We begin by running the regression:

$Tbill_t = a_0 + a_1R5_t + a_2R10_t$
```{r}
# Run regression from book
t_dep_fit <- lm(tbill ~ r5 + r10, data = df)
summary(t_dep_fit)
residuals <- t_dep_fit$residuals
```
The estimates differ from those in the book. However, when we estimated the same regression using the data used in the book, we obtianed similar values. We proceed by testing if the residuals are serially conintegrated using ADF-tests.

We plot the residual series:
```{r}
df <- df %>% mutate(residuals = residuals)
df %>% ggplot(aes(x = DATE, y = residuals)) +
  geom_line() +
  labs(title = "Could be stationary")
```



Next, we perform ADF test on residuals (i.e. Engle Granger). We select lag length using the AIC criteria, with a maximal lag length of 10.
```{r}
adf_res <- ur.df(residuals, type = "none", lags = 10, selectlags = "AIC")
summary(adf_res) 
```
We need Engle - Granger critical values to evaluate the test statistic,`r round(adf_res@teststat, 3)` against -3.785. The test statistic is small enough for us to reject the null of the series having a unit root, hence the series is stationary and there is cointegration between the variables.

Next, we test whether the residuals follow a white noise process. We use the Ljung box test with a lag length of 9 (the optimal number by the AIC in the ADF test).

```{r}
Box.test(residuals, lag = 9, type = "Ljung-Box")
```
Given the low p-value, we reject the null hypothesis of independence, meaning the residuals are correlated. Anyway, we move on.

## 4c, similar to above
```{r}
# Run rgression from book
t_dep_fit <- lm(r10 ~ tbill + r5, data = df)
summary(t_dep_fit)
residuals <- t_dep_fit$residuals
```
```{r}
df <- df %>% mutate(residuals = residuals)
df %>% ggplot(aes(x = DATE, y = residuals)) +
  geom_line() +
  labs(title = "Could be stationary", y = "R10, residuals")
```

```{r}
adf_res <- ur.df(residuals, type = "none", lags = 10, selectlags = "AIC")
summary(adf_res) 
```
A test statistic value of `r adf_res@teststat` is obtained and evaluated against -3.785, and we are not able to reject the null of a unit root. This means we can't draw the conclusion that the three variables are cointegrated form this test. This contradicts what we found earlier and indicates that the Engle Granger test might not be the most suitable test for cointegration.

## 4d
Use the Johansen procedure
```{r}
# Function to use: ca.jo from the urca package
# This link might be helpful:
#https://www.quantstart.com/articles/Johansen-Test-for-Cointegrating-Time-Series-Analysis-in-R

tbill <- ts(df$tbill)
r5 <- ts(df$r5)
r10 <- ts(df$r10)
ts <- window(ts.union(tbill, r5, r10))

summary(ca.jo(ts, type = "trace", ecdet = "const", K = 7))
```
i) By the trace test, we can reject that there are 0 distinct cointegrating vectors, meaning there is some cointegration. However, we can't reject that there are less or equal to 1 cointegrating relationships, so we conclude that there is only one cointegrating relationship present in the data.
```{r}
summary(ca.jo(ts, type = "eigen", ecdet = "const", K = 7))

```

ii) The observed test statistics support the notion that there is one cointegrating relationship present in the data.

We don't verify that the cointrgrated vector is as was stated in the book since the data has changed. In theory, we'd obtain a zero vector after applying the constants to the data, but we don't have the proper constants.


## 4e
Check to determine whether the individual interest pairs are cointegrated. In particular, is R5 with cointegrated with R10.

```{r}
# Very similar to what we did in b & c, just with one variable less.
# Run rgression from book
t_dep_fit <- lm(r5 ~ r10, data = df)
summary(t_dep_fit)
residuals <- t_dep_fit$residuals
```

```{r}
df <- df %>% mutate(residuals = residuals)
df %>% ggplot(aes(x = DATE, y = residuals)) +
  geom_line() +
  labs(title = "Could be stationary")
```
```{r}
adf_res <- ur.df(residuals, type = "none", lags = 10, selectlags = "AIC")
summary(adf_res) 
```
The observed value of the test statistic is `r adf_res@teststat` which is almost small enough to reject the null of no cointegrating relationship at the 5% level.
